UD651 Lesson 2 Notes by mrandrewandrade
========================================================

UD651 Udacity's Exploratory Data Analysis Course Notes by Andrew Andrade || mrandrewandrade.com  || March 22nd, 2014 || (CC BY-NC 2.5)
***

### Install R and RStudio

Download and install the R programming language FIRST at http://cran.rstudio.com. 

After you install R, you can download and install RStudio from http://www.rstudio.com.

### RStudio Layout
Layout Guide:
http://dss.princeton.edu/training/RStudio101.pdf

***

## Demystifying R Key Notes
Learn to run nine commands: c, nchar, data, str, dim, names, row.names, head, and tail.

### Part 1: Getting Started


```{r}
# Store instructor names in a vector
udacious_instructors <- c("Chris Saden", "Lauren Castellano",
              "Sarah Spikes","Dean Eckles",
              "Andy Brown", "Moira Burke",
              "Kunal Chawla")

# Display contents of vector
udacious_instructors

# Store 1 to 10 then 1 to 20 in a vector
numbers <- c(1:10)

#display
numbers

#append 11:20
numbers <- c(numbers, 11:20)

#display
numbers
```


#### Display number of characters using nchar:
```{r}
udacious <- c("Chris Saden", "Lauren Castellano",
              "Sarah Spikes","Dean Eckles",
              "Andy Brown", "Moira Burke",
              "Kunal Chawla", "Andrew Andrade")

# vector that contains the number of characters
# for each of the names in udacious, including my name.
nchar_output = nchar(udacious)
nchar_output

# Check if name in vector contains exactly 11 characters
nchar_output == 11

# Output names with 11 charaters only
udacious[nchar_output  == 11]

```

### Example 2: Understanding Dataframes

```{r}
#load mtcars data
data(mtcars)
#display the variable names in mtcars
names(mtcars)
#display help
?mtcars
#display data frame in table form
mtcars

#display structur of data fram
str(mtcars)

#displays the size of the data frame
dim(mtcars)

#help for row names
?row.names

# Change the row names of the cars to numbers.
row.names(mtcars) <- c(1:32)
# Show changes:
mtcars

#reload the data to reset row names
data(mtcars)

#display first 6 entries (default)
head(mtcars)

#display first 10 entries
head(mtcars, 10)

#display last 3 entries
tail(mtcars, 3)

```

#### Note: R uses 1 to start indexing (AND NOT ZERO BASED INDEXING as is true of many other programming languages.)



### Example 3: Examine Data More Closely


```{r}
#Access an individual variable (or column) from the data frame
mtcars$mpg
#Find average of that column
mean(mtcars$mpg)
```


***